---
title: "PA1_template.Rmd"
author: "Zahra Mohammadi"
date: "3/30/2020"
output: html_document
---
## Reproducible Research: Peer Assessment 1

#### Set echo = TRUE for all chuncks to show R codes in markdown

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

#### Removing hashtags in final result

```{r include=TRUE}
knitr::opts_chunk$set(comment = NA)
```

#### Setting working directory
```{r}
setwd('C:\\Users\\moham\\Documents\\R')
```

#### Loading and Preprocessing the data

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip")

#### Unziping the data

```{r}
if (!file.exists('activity.csv')) {
  unzip(zipfile = "activity.zip")
}
```
#### Reading the data
```{r} 
activityData <- read.csv(file="activity.csv", header=TRUE)
```

#### What is the mean total number of steps taken per day?

##### Calculate the total steps taken per day

```{r} 
totalSteps <- aggregate(steps ~ date, activityData, sum, na.rm = TRUE) 
```

#### Make a histogram of the total number of steps taken per day

```{r}
hist(totalSteps$steps,
     main = "Total Steps per Day",
     xlab = "Number of Steps", breaks = 15)
```

#### Calculate and report the mean and median of total steps taken per day

```{r}
meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medSteps <- median(totalSteps$steps, na.rm = TRUE)
paste("The mean total steps in a day is", meanSteps, "and the median is", medSteps, ".")

```

#### What is the average daily activity pattern?
#### Make a time-series plot of the 5-minute interval and the average number of steps taken, averaged acoss all days.

```{r}
library(ggplot2)
meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Which 5-minute interval across all days contain the maximum number of steps
```{r}
maxInt <- meanStepsByInt[which.max(meanStepsByInt$steps),]
paste("The interval with the largest average number of steps is interval", maxInt$interval, ".")
```
#### Imputing missing values 
#### Calculate and report the total number of missing values in the dataset
```{r}
missingVals <- is.na(activityData$steps)
sum(missingVals)
```

#### Devise a strategy for filling in all of the missing values
#### Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
imp_activityData <- 
transform(activityData, steps = ifelse(is.na(activityData$steps),meanStepsByInt$steps[match(activityData$interval, meanStepsByInt$interval)],activityData$steps))
```

#### Make a histogram of the total number of steps taken each day and
#### and report the mean and median.
```{r}
impStepsByInt <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impStepsByInt$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")
impMeanSteps <- mean(impStepsByInt$steps, na.rm = TRUE)
impMedSteps <- median(impStepsByInt$steps, na.rm = TRUE)
diffMean = impMeanSteps - meanSteps
diffMed = impMedSteps - medSteps
diffTotal = sum(impStepsByInt$steps) - sum(totalSteps$steps)
paste("The new mean total steps in a day is", impMeanSteps, "and the median is", impMedSteps, ", instead of", meanSteps, "and", medSteps, "respectively.")
```

#### Are there differences in activity patterns between weekdays and weekends?
#### Create a new factor variable in the dataset with two levels - "weekend" and "weekday"
```{r}
DayType <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
    return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
    return ("weekend")
  else
    stop ("Invalid Date Format.")
}
imp_activityData$date <- as.Date(imp_activityData$date)
imp_activityData$day <- sapply(imp_activityData$date, FUN = DayType)
```

#### Make a panel plot containnig a time-series plot of the 5-minute interval
#### and the average number of steps taken across all weekdays or weekends
```{r}
meanStepsByDay <- aggregate(steps ~ interval + day, imp_activityData, mean)
ggplot(data = meanStepsByDay, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))

```









